---
title: 'Task 2: Willamette Falls fish passage time series summary'
author: "Claire Meuter"
date: "2023-03-19"
output: 
  html_document:
    theme: journal
    code_folding: hide
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(lubridate)
```
## Overview {.tabset}
-An engaging image (with caption, incl. photo credit) that is relevant to the dataset
-A brief summary (3 - 4 sentences) of the dataset, and what is included in this report
-A map of the fish ladder location (you can make this in R on your own, or include an existing map appropriately licensed, with attribution)
-A professionally formatted data citation

### Original Time Series 
```{r}
# read in the data and usual clean up 
fish <- read_csv(here("data","willamette_fish_passage.csv")) %>% 
  janitor::clean_names() %>% 
  select(date, steelhead, coho, jack_coho) %>% #selecting for the fish of interest 
  replace(is.na(.),0) %>%  #replacing nas with 0 
  mutate(date = lubridate::mdy(date)) #making my date column class: date 

  
#Now I need to covert the df to tsibble (a times series df )
fish_ts <- fish %>% 
  as_tibble(key = NULL, index = date)

# pivot data so species is in same column 
fish_ts_pivot <- fish_ts %>% 
  pivot_longer(cols = 2:4, names_to = "species", values_to = "count")


## Now i can visualize the data 
ggplot(data = fish_ts_pivot, aes(x = date, y = count, color = species)) +
  geom_line() + 
  facet_wrap(~species, nrow = 3)

```


### tab 2
(content)

### tab 3 
(content)
